The Unicode Cookbook \linebreak for Linguists

### 2018-06-23 book/main.tex
ready for release, except doi

You can also install the \texttt{segments} package from the GitHub repository:\footnote{\url{https://github.com/cldf/segments}}

\begin{lstlisting}[language=bash]
  $ git clone https://github.com/cldf/segments
  $ cd segments
  $ python setup.py develop
\end{lstlisting}

\subsection*{Application programming interface}
The \texttt{segments} API can be accessed by importing the package into Python or by writing a Python script. Here is an example of how to import the libraries, create a tokenizer object, tokenize a string, and create an orthography profile.

\begin{lstlisting}
>>> from segments.tokenizer import Tokenizer
\end{lstlisting}

\noindent The \texttt{characters} function will segment a string at Unicode code points.

% \lstset{extendedchars=false, escapeinside=**}
\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t = Tokenizer()
>>> result = t.characters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@c ̂ h a ́ ɾ a ̃ ̌ c t ʼ ɛ ↗ ʐ ː | # k ͡ p@*)'
\end{lstlisting}

\noindent The \texttt{grapheme\_clusters} function will segment text at the Unicode Extended Grapheme Cluster boundaries.\footnote{\url{http://www.unicode.org/reports/tr18/tr18-19.html\#Default_Grapheme_Clusters}}

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t.grapheme_clusters('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

\noindent The \texttt{grapheme\_clusters} function is the default segmentation algorithm for the \texttt{segments.Tokenizer}. It is useful when you encounter a text that you want to tokenize to identify orthographic or transcription elements.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | # k͡ p@*)'
\end{lstlisting}

\noindent The \texttt{ipa} parameter forces grapheme segmentation for IPA strings.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', ipa=True)
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c tʼ ɛ ↗ ʐː | # k͡p@*)'
\end{lstlisting}

\noindent We can also load an orthography profile and tokenize an input string with it. In the data directory, we've placed an example orthography profile. Let's have a look at it using \texttt{more} on the command line.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
  $ more data/orthography-profile.tsv
  Grapheme  IPA   XSAMPA  COMMENT
  a         a     a
  aa        (*@aː@*)    (*@a:@*)
  b         b     b
  c         c     c
  ch        (*@tʃ@*)    tS
  -         NULL  NULL    "comment with   tab"
  on        (*@õ@*)     o~
  n         n     n
  ih        (*@í@*)     i_H
  inh       (*@ĩ́@*)     i~_H
\end{lstlisting}

\noindent An orthography profile is a delimited UTF-8 text file (here we use tab as a delimiter for reading ease). The first column must be labelled \texttt{Grapheme}. Each row in the \texttt{Grapheme} column specifies graphemes that may be found in the orthography of the input text. In this example, we provide additional columns IPA and XSAMPA, which are mappings from our graphemes to their IPA and XSAMPA transliterations. The final column \texttt{COMMENT} is for comments; if you want to use a tab ``quote that     string''!

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t('(*@aabchonn-ih@*)')
>>> '(*@aa b ch on n - ih@*)'
\end{lstlisting}

\noindent This example shows how we can tokenize input text into our orthographic specification. We can also segment graphemes and transliterate them into other formats, which is useful when you have sources with different orthographies, but you want to be able to compare them using a single representation like IPA.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}]
>>> t.transform('(*@aabchonn-ih@*)', 'IPA')
>>> '(*@aː b tʃ õ n í@*)'
\end{lstlisting}

@@

\noindent And we can transliterate to XSAMPA.

\begin{lstlisting}[language=bash]
  $ cat sources/german.txt | segments --mapping=XSAMPA 
	--profile=data/german-orthography-profile.tsv tokenize

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ú & 1 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Ù & 1 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
  Û & 1 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 1 (default settings, splitting grapheme clusters)} 
\label{tab:profile1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 1 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 1 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 1 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 3 & U+0055 & LATIN CAPITAL LETTER U \\ 
  Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 2 (splitting by code points)} 
\label{tab:profile2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 3 (splitting by NFC code points)} 
\label{tab:profile3}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 2 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 2 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 2 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 6 & U+0055 & LATIN CAPITAL LETTER U \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 4 (splitting by NFD code points)} 
\label{tab:profile4}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 2 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 2 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 5 (splitting by graphemes after NFD)} 
\label{tab:profile5}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 6 (splitting by graphemes after NFC)} 
\label{tab:profile6}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
sh &  &  &  \\ 
  ch &  &  &  \\ 
  sch &  &  &  \\ 
  ie &  &  &  \\ 
  oo &  &  &  \\ 
  a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
  c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
  e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
  f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
  h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
  i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
  m & 4 & U+006D & LATIN SMALL LETTER M \\ 
  o & 2 & U+006F & LATIN SMALL LETTER O \\ 
  p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
  s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Manually edited profile skeleton} 
\label{tab:profile_skeleton1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
mi & s &  &  &  \\ 
   & sh &  &  &  \\ 
   & ch &  &  &  \\ 
   & sch &  &  &  \\ 
   & ie &  &  &  \\ 
   & oo &  &  &  \\ 
   & a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
   & c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
   & e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
   & f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
   & h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
   & i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
   & m & 4 & U+006D & LATIN SMALL LETTER M \\ 
   & o & 2 & U+006F & LATIN SMALL LETTER O \\ 
   & p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
   & s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with contextual specification for <s>} 
\label{tab:profile_skeleton2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & Replacement \\ 
  \midrule
 & a &  &  & a \\ 
   & c &  &  & c \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   & n &  &  & n \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with empty columns for editing contexts} 
\label{tab:profile_editing_1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & [ie] &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with regex as context} 
\label{tab:profile_editing_2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Thu Jun 21 15:02:19 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & Vfront &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  & Vfront & e \\ 
   & i &  & Vfront & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with Class as context} 
\label{tab:profile_editing_3}
\end{table}

>>>>

You can also install the \texttt{segments} package from the GitHub repository,\footnote{\url{https://github.com/cldf/segments}} in particular if you would like to contribute to the code base:\footnote{\url{https://github.com/cldf/segments/blob/master/CONTRIBUTING.md}}

\begin{lstlisting}[language=bash, basicstyle=\myfont]
  $ git clone https://github.com/cldf/segments
  $ cd segments
  $ python setup.py develop
\end{lstlisting}

\subsection*{Application programming interface}
The \texttt{segments} API can be accessed by importing the package into Python. Here is an example of how to import the library, create a tokenizer object, tokenize a string, and create an orthography profile. Begin by importing the \texttt{Tokenizer} from the \texttt{segments} library.

\begin{lstlisting}[basicstyle=\myfont]
>>> from segments.tokenizer import Tokenizer
\end{lstlisting}

\noindent Next, instantiate a tokenizer object, which takes optional arguments for an orthography profile and an orthography profile rules file.

\begin{lstlisting}[basicstyle=\myfont]
>>> t = Tokenizer()
\end{lstlisting}

\noindent The default tokenization strategy is to segment some input text at the Unicode Extended Grapheme Cluster boundaries,\footnote{\url{http://www.unicode.org/reports/tr18/tr18-19.html\#Default_Grapheme_Clusters}} and to return, by default, a space-delimited string of graphemes. White space between input string sequences is by default separated by a hash symbol <\#>, which is a linguistic convention used to denote word boundaries. The default grapheme tokenization is useful when you encounter a text that you want to tokenize to identify potential orthographic or transcription elements.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)')
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | \# k͡ p@*)'
\end{lstlisting}

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', segment_separator='(*@-@*)')
>>> print(result)
>>> '(*@ĉ-h-á-ɾ-ã̌-c-t-ʼ-ɛ-↗-ʐ-ː-| \# k͡ -p@*)'
\end{lstlisting}

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont, showstringspaces=false]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', separator=' // '))
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐ ː | // k͡ p@*)'
\end{lstlisting}

\noindent The optional \texttt{ipa} parameter forces grapheme segmentation for IPA strings.\footnote{\url{https://en.wikipedia.org/wiki/International\_Phonetic\_Alphabet}} Note here that Unicode Spacing Modifier Letters,\footnote{\url{https://en.wikipedia.org/wiki/Spacing\_Modifier\_Letters}} such as <ː> and <\dia{0361}{\large\fontspec{CharisSIL}◌}>, will be segmented together with base characters (although you might need orthography profiles and rules to correct these in your input source; see Section \ref{pitfall-different-notions-of-diacritics} for details).

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> result = t('(*@ĉháɾã̌ctʼɛ↗ʐː| k͡p@*)', ipa=True)
>>> print(result)
>>> '(*@ĉ h á ɾ ã̌ c t ʼ ɛ ↗ ʐː | \# k͡p@*)'
\end{lstlisting}

\noindent You can also load an orthography profile and tokenize input strings with it. In the data directory,\footnote{https://github.com/unicode-cookbook/recipes/tree/master/Basics/data} we've placed an example orthography profile. Let's have a look at it using \texttt{more} on the command line.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont, showstringspaces=false]
  $ more data/orthography-profile.tsv
  Grapheme  IPA   XSAMPA  COMMENT
  a         a     a
  aa        (*@aː@*)    (*@a:@*)
  b         b     b
  c         c     c
  ch        (*@tʃ@*)    tS
  (*@-@*)         NULL  NULL    "comment with   tab"
  on        (*@õ@*)     o~
  n         n     n
  ih        (*@í@*)     i_H
  inh       (*@ĩ́@*)     i~_H
\end{lstlisting}

\noindent An orthography profile is a delimited UTF-8 text file (here we use tab as a delimiter for reading ease). The first column must be labeled \texttt{Grapheme}, as discussed in Section \ref{formal-specification-of-orthography-profiles}. Each row in the \texttt{Grapheme} column specifies graphemes that may be found in the orthography of the input text. In this example, we provide additional columns \texttt{IPA} and \texttt{XSAMPA}, which are mappings from our graphemes to their IPA and X-SAMPA transliterations. The final column \texttt{COMMENT} is for comments; if you want to use a tab ``quote that     string''!

@@

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> t('(*@aabchonn-ih@*)')
>>> '(*@aa b ch on n - ih@*)'
\end{lstlisting}

\noindent This example shows how we can tokenize input text into our orthographic specification. We can also segment graphemes and transliterate them into other forms, which is useful when you have sources with different orthographies, but you want to be able to compare them using a single representation like IPA or X-SAMPA.

\begin{lstlisting}[extendedchars=false, escapeinside={(*@}{@*)}, basicstyle=\myfont]
>>> t('(*@aabchonn-ih@*)', column='IPA')
>>> '(*@aː b tʃ õ n í@*)'
\end{lstlisting}

@@

\noindent And we can transliterate to X-SAMPA.

\begin{lstlisting}[language=bash, basicstyle=\myfont, extendedchars=false, escapeinside={(*@}{@*)}]
  $ cat sources/german.txt | segments (*@--@*)mapping=XSAMPA 
	(*@--@*)profile=data/german(*@-@*)orthography(*@-@*)profile.tsv tokenize

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:13 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ú & 1 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Ù & 1 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
  Û & 1 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 1 (default settings, splitting grapheme clusters)} 
\label{tab:profile1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 1 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 1 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 1 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 3 & U+0055 & LATIN CAPITAL LETTER U \\ 
  Ú & 1 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 1 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 1 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 2 (splitting by code points)} 
\label{tab:profile2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 3 (splitting by NFC code points)} 
\label{tab:profile3}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & Unicode Name \\ 
  \midrule
́ & 2 & U+0301 & COMBINING ACUTE ACCENT \\ 
  ̀ & 2 & U+0300 & COMBINING GRAVE ACCENT \\ 
  ̂ & 2 & U+0302 & COMBINING CIRCUMFLEX ACCENT \\ 
  U & 6 & U+0055 & LATIN CAPITAL LETTER U \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 4 (splitting by NFD code points)} 
\label{tab:profile4}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+0055, U+0301 & LATIN CAPITAL LETTER U, COMBINING ACUTE ACCENT \\ 
  Ù & 2 & U+0055, U+0300 & LATIN CAPITAL LETTER U, COMBINING GRAVE ACCENT \\ 
  Û & 2 & U+0055, U+0302 & LATIN CAPITAL LETTER U, COMBINING CIRCUMFLEX ACCENT \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 5 (splitting by graphemes after NFD)} 
\label{tab:profile5}
\end{table}

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[H]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Gr. & Freq. & Codepoint & Unicode Name \\ 
  \midrule
Ú & 2 & U+00DA & LATIN CAPITAL LETTER U WITH ACUTE \\ 
  Ù & 2 & U+00D9 & LATIN CAPITAL LETTER U WITH GRAVE \\ 
  Û & 2 & U+00DB & LATIN CAPITAL LETTER U WITH CIRCUMFLEX \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Profile 6 (splitting by graphemes after NFC)} 
\label{tab:profile6}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{llll}
  \toprule
Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
sh &  &  &  \\ 
  ch &  &  &  \\ 
  sch &  &  &  \\ 
  ie &  &  &  \\ 
  oo &  &  &  \\ 
  a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
  c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
  e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
  f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
  h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
  i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
  m & 4 & U+006D & LATIN SMALL LETTER M \\ 
  o & 2 & U+006F & LATIN SMALL LETTER O \\ 
  p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
  s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Manually edited profile skeleton} 
\label{tab:profile_skeleton1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Frequency & Codepoint & UnicodeName \\ 
  \midrule
mi & s &  &  &  \\ 
   & sh &  &  &  \\ 
   & ch &  &  &  \\ 
   & sch &  &  &  \\ 
   & ie &  &  &  \\ 
   & oo &  &  &  \\ 
   & a & 2 & U+0061 & LATIN SMALL LETTER A \\ 
   & c & 2 & U+0063 & LATIN SMALL LETTER C \\ 
   & e & 3 & U+0065 & LATIN SMALL LETTER E \\ 
   & f & 1 & U+0066 & LATIN SMALL LETTER F \\ 
   & h & 4 & U+0068 & LATIN SMALL LETTER H \\ 
   & i & 3 & U+0069 & LATIN SMALL LETTER I \\ 
   & m & 4 & U+006D & LATIN SMALL LETTER M \\ 
   & o & 2 & U+006F & LATIN SMALL LETTER O \\ 
   & p & 2 & U+0070 & LATIN SMALL LETTER P \\ 
   & s & 4 & U+0073 & LATIN SMALL LETTER S \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with contextual specification for <s>} 
\label{tab:profile_skeleton2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & Replacement \\ 
  \midrule
 & a &  &  & a \\ 
   & c &  &  & c \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   & n &  &  & n \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with empty columns for editing contexts} 
\label{tab:profile_editing_1}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & [ie] &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  &  & e \\ 
   & i &  &  & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with regex as context} 
\label{tab:profile_editing_2}
\end{table}

@@

% latex table generated in R 3.5.0 by xtable 1.8-2 package
% Sat Jun 23 12:54:14 2018
\begin{table}[htb]
\centering
\begingroup\scriptsize
\begin{tabular}{lllll}
  \toprule
Left & Grapheme & Right & Class & IPA \\ 
  \midrule
 & c & Vfront &  & tʃ \\ 
   & a &  &  & a \\ 
   & n &  &  & n \\ 
   & c &  &  & k \\ 
   & e &  & Vfront & e \\ 
   & i &  & Vfront & i \\ 
   \bottomrule
\end{tabular}
\endgroup
\caption{Orthography profile with Class as context} 
\label{tab:profile_editing_3}
\end{table}
